Verification of the CVM algorithm with a Functional Probabilistic Invariant

### 2025-02-25 paper/sec/formalization_neg_dep.tex
Paper: Further edits on neg assoc section   * Add footnote about the proof of perm distributions.   * Due to space reasons: Replace example snippet (Property 1) with Hoeffding ineq for neg assoc vars, which requires less explanation.

Let us summarize a few key formalized results for negatively associated random variables used in our CVM proof:

@@

\ neg{\isacharunderscore}{\kern0pt}assoc{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}mult{\isacharunderscore}{\kern0pt}mono{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{fixes}\ f\ g\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharprime}{\kern0pt}i\ {\isasymRightarrow}\ {\isacharprime}{\kern0pt}c{\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}linorder{\isacharunderscore}{\kern0pt}topology{\isacharparenright}{\kern0pt}\ {\isasymRightarrow}\ real{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}neg{\isacharunderscore}{\kern0pt}assoc\ X\ I{\isacartoucheclose}\ {\isacartoucheopen}J\ {\isasymsubseteq}\ I{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}square{\isacharunderscore}{\kern0pt}integrable\ M\ {\isacharparenleft}{\kern0pt}f\ {\isasymcirc}\ flip\ X{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}square{\isacharunderscore}{\kern0pt}integrable\ M\ {\isacharparenleft}{\kern0pt}g\ {\isasymcirc}\ flip\ X{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}monotone\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isasymge}\isactrlbsub {\isasymeta}\isactrlesub {\isacharparenright}{\kern0pt}\ f{\isacartoucheclose}\ {\isacartoucheopen}monotone\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isasymge}\isactrlbsub {\isasymeta}\isactrlesub {\isacharparenright}{\kern0pt}\ g{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}depends{\isacharunderscore}{\kern0pt}on\ f\ J{\isacartoucheclose}\ {\isacartoucheopen}depends{\isacharunderscore}{\kern0pt}on\ g\ {\isacharparenleft}{\kern0pt}I{\isacharminus}{\kern0pt}J{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}f\ {\isasymin}\ borel{\isacharunderscore}{\kern0pt}measurable\ {\isacharparenleft}{\kern0pt}Pi\isactrlsub M\ J\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ borel{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}g\ {\isasymin}\ borel{\isacharunderscore}{\kern0pt}measurable\ {\isacharparenleft}{\kern0pt}Pi\isactrlsub M\ {\isacharparenleft}{\kern0pt}I{\isacharminus}{\kern0pt}J{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ borel{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isasymintegral}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}{\isacharasterisk}{\kern0pt}g{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}\ {\isasympartial}M{\isacharparenright}{\kern0pt}{\isasymle}{\isacharparenleft}{\kern0pt}{\isasymintegral}\isasymomega{\isachardot}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ \isasymomega{\isacharparenright}{\kern0pt}{\isasympartial}M{\isacharparenright}{\kern0pt}{\isacharasterisk}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymintegral}{\kern0pt}\isasymomega{\isachardot}{\kern0pt}\ g{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ \isasymomega{\isacharparenright}{\kern0pt}{\isasympartial}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}

@@

This corresponds to Property~\ref{it:neg_dep_props:mult_mono} of \cref{pro:neg_dep_props}:
If $X_i$ for $i \in I$ form a set of negatively associated variables, and $f$, $g$ are simultaneously monotone (or simultaneously antimonotone), real-valued, square-integrable, measurable functions, depending on disjoint components of $I$ then the expectations of the product are smaller than then product of the expectations of the composition of the functions $f,g$ with random variables $X_i$.
Note that, if $X_i$ where independent the above would be true, without the monotonicity assumption.

The following is the key result we use for the verification of the new CVM algorithm:
The negative-associativity of the indicator functions of random $k$-subsets of a finite set $S$ (with cardinality greater than $k$).
\begin{isabelle_cm}
\isacommand{lemma}\isamarkupfalse%
\ n{\isacharunderscore}{\kern0pt}subsets{\isacharunderscore}{\kern0pt}distribution{\isacharunderscore}{\kern0pt}neg{\isacharunderscore}{\kern0pt}assoc{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}finite\ S{\isacartoucheclose}\ {\isacartoucheopen}k\ {\isasymle}\ card\ S{\isacartoucheclose}\isanewline
\ \ \isakeyword{defines}\ {\isacartoucheopen}p\ {\isasymequiv}\ pmf{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}set\ {\isacharbraceleft}{\kern0pt}T{\isachardot}{\kern0pt}\ T\ {\isasymsubseteq}\ S\ {\isasymand}\ card\ T\ {\isacharequal}{\kern0pt}\ k{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}measure{\isacharunderscore}{\kern0pt}pmf{\isachardot}{\kern0pt}neg{\isacharunderscore}{\kern0pt}assoc\ p\ {\isacharparenleft}{\kern0pt}{\isasymin}{\isacharparenright}{\kern0pt}\ S{\isacartoucheclose}
\end{isabelle_cm}

%TODO: I suggest to put n_subsets_distribution_neg_assoc instead here instead, since the Hoeffding inequality is never used in our CVM proof.
%The following is the well-known Hoeffding inequality~\cite{hoeffding1963} for negatively associated random variables:
%\begin{isabelle_cm}
%\isacommand{lemma}\isamarkupfalse%
%\ hoeffding{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}two{\isacharunderscore}{\kern0pt}sided{\isacharcolon}{\kern0pt}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}neg{\isacharunderscore}{\kern0pt}assoc\ X\ I{\isacartoucheclose}\ {\isacartoucheopen}finite\ I{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ a\ i\ {\isasymle}\ b\ i{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ AE\ {\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}a\ i{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}b\ i{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}I\ {\isasymnoteq}\ {\isacharbraceleft}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{defines}\ {\isacartoucheopen}n\ {\isasymequiv}\ real\ {\isacharparenleft}{\kern0pt}card\ I{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{defines}\ {\isacartoucheopen}{\isasymmu}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ expectation\ {\isacharparenleft}{\kern0pt}X\ i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymdelta}\ {\isasymge}\ {\isadigit{0}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isachargreater}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{shows}\ {\isacartoucheopen}{\isasymP}{\isacharparenleft}{\kern0pt}{\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ {\isasymbar}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isasymmu}{\isasymbar}\ {\isasymge}\ {\isasymdelta}{\isacharasterisk}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharasterisk}{\kern0pt}exp\ {\isacharparenleft}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{2}}{\isacharasterisk}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharasterisk}{\kern0pt}{\isasymdelta}{\isacharparenright}{\kern0pt}\isactrlsup {\isadigit{2}}\ {\isacharslash}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\end{isabelle_cm}

%TODO: I think the following footnote should be placed here, if we put negative association for permutation distribution here.
%\footnote{We could not verify the original proof by Joag-Dev and Proschan, which seems to be incorrect.
%However Dubashi presented a correct proof later using the FKG inequality~\cite[Th. 10]{dubhashi1996}.}

>>>>

Let us summarize a few key formalized results for negatively associated random variables:

@@

\ hoeffding{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}two{\isacharunderscore}{\kern0pt}sided{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}neg{\isacharunderscore}{\kern0pt}assoc\ X\ I{\isacartoucheclose}\ {\isacartoucheopen}finite\ I{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ a\ i\ {\isasymle}\ b\ i{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ AE\ {\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}a\ i{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}b\ i{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}I\ {\isasymnoteq}\ {\isacharbraceleft}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{defines}\ {\isacartoucheopen}n\ {\isasymequiv}\ real\ {\isacharparenleft}{\kern0pt}card\ I{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{defines}\ {\isacartoucheopen}{\isasymmu}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ expectation\ {\isacharparenleft}{\kern0pt}X\ i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymdelta}\ {\isasymge}\ {\isadigit{0}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isachargreater}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}{\isasymP}{\isacharparenleft}{\kern0pt}{\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ {\isasymbar}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isasymmu}{\isasymbar}\ {\isasymge}\ {\isasymdelta}{\isacharasterisk}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharasterisk}{\kern0pt}exp\ {\isacharparenleft}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{2}}{\isacharasterisk}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharasterisk}{\kern0pt}{\isasymdelta}{\isacharparenright}{\kern0pt}\isactrlsup {\isadigit{2}}\ {\isacharslash}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}

@@

The above is the well-known Hoeffding inequality~\cite{hoeffding1963} for negatively associated random variables.
%\begin{isabelle_cm}
%\isacommand{lemma}\isamarkupfalse%
%\ neg{\isacharunderscore}{\kern0pt}assoc{\isacharunderscore}{\kern0pt}imp{\isacharunderscore}{\kern0pt}mult{\isacharunderscore}{\kern0pt}mono{\isacharcolon}{\kern0pt}\isanewline
%\ \ \isakeyword{fixes}\ f\ g\ {\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isacharprime}{\kern0pt}i\ {\isasymRightarrow}\ {\isacharprime}{\kern0pt}c{\isacharcolon}{\kern0pt}{\isacharcolon}{\kern0pt}linorder{\isacharunderscore}{\kern0pt}topology{\isacharparenright}{\kern0pt}\ {\isasymRightarrow}\ real{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}neg{\isacharunderscore}{\kern0pt}assoc\ X\ I{\isacartoucheclose}\ {\isacartoucheopen}J\ {\isasymsubseteq}\ I{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}square{\isacharunderscore}{\kern0pt}integrable\ M\ {\isacharparenleft}{\kern0pt}f\ {\isasymcirc}\ flip\ X{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}square{\isacharunderscore}{\kern0pt}integrable\ M\ {\isacharparenleft}{\kern0pt}g\ {\isasymcirc}\ flip\ X{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}monotone\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isasymge}\isactrlbsub {\isasymeta}\isactrlesub {\isacharparenright}{\kern0pt}\ f{\isacartoucheclose}\ {\isacartoucheopen}monotone\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymle}{\isasymge}\isactrlbsub {\isasymeta}\isactrlesub {\isacharparenright}{\kern0pt}\ g{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}depends{\isacharunderscore}{\kern0pt}on\ f\ J{\isacartoucheclose}\ {\isacartoucheopen}depends{\isacharunderscore}{\kern0pt}on\ g\ {\isacharparenleft}{\kern0pt}I{\isacharminus}{\kern0pt}J{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}f\ {\isasymin}\ borel{\isacharunderscore}{\kern0pt}measurable\ {\isacharparenleft}{\kern0pt}Pi\isactrlsub M\ J\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ borel{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\ \isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}g\ {\isasymin}\ borel{\isacharunderscore}{\kern0pt}measurable\ {\isacharparenleft}{\kern0pt}Pi\isactrlsub M\ {\isacharparenleft}{\kern0pt}I{\isacharminus}{\kern0pt}J{\isacharparenright}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymlambda}{\isacharunderscore}{\kern0pt}{\isachardot}{\kern0pt}\ borel{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{shows}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isasymintegral}{\kern0pt}{\isasymomega}{\isachardot}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}{\isacharasterisk}{\kern0pt}g{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}\ {\isasympartial}M{\isacharparenright}{\kern0pt}{\isasymle}{\isacharparenleft}{\kern0pt}{\isasymintegral}\isasymomega{\isachardot}{\kern0pt}\ f{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ \isasymomega{\isacharparenright}{\kern0pt}{\isasympartial}M{\isacharparenright}{\kern0pt}{\isacharasterisk}{\kern0pt}{\isacharparenleft}{\kern0pt}{\isasymintegral}{\kern0pt}\isasymomega{\isachardot}{\kern0pt}\ g{\isacharparenleft}{\kern0pt}{\isasymlambda}i{\isachardot}{\kern0pt}\ X\ i\ \isasymomega{\isacharparenright}{\kern0pt}{\isasympartial}M{\isacharparenright}{\kern0pt}{\isacartoucheclose}
%\end{isabelle_cm}
%This corresponds to Property~\ref{it:neg_dep_props:mult_mono} of \cref{pro:neg_dep_props}:
%If $X_i$ for $i \in I$ form a set of negatively associated variables, and $f$, $g$ are simultaneously monotone (or simultaneously antimonotone), real-valued, square-integrable, measurable functions, depending on disjoint components of $I$ then the expectations of the product are smaller than then product of the expectations of the composition of the functions $f,g$ with random variables $X_i$.
%Note that, if $X_i$ where independent the above would be true, without the monotonicity assumption.
The following is the key result we use for the verification of the new CVM algorithm:
The negative-associativity of the indicator functions of random $k$-subsets of a finite set $S$ (with cardinality greater than $k$).
\ n{\isacharunderscore}{\kern0pt}subsets{\isacharunderscore}{\kern0pt}distribution{\isacharunderscore}{\kern0pt}neg{\isacharunderscore}{\kern0pt}assoc{\isacharcolon}{\kern0pt}\isanewline
\ \ \isakeyword{assumes}\ {\isacartoucheopen}finite\ S{\isacartoucheclose}\ {\isacartoucheopen}k\ {\isasymle}\ card\ S{\isacartoucheclose}\isanewline
\ \ \isakeyword{defines}\ {\isacartoucheopen}p\ {\isasymequiv}\ pmf{\isacharunderscore}{\kern0pt}of{\isacharunderscore}{\kern0pt}set\ {\isacharbraceleft}{\kern0pt}T{\isachardot}{\kern0pt}\ T\ {\isasymsubseteq}\ S\ {\isasymand}\ card\ T\ {\isacharequal}{\kern0pt}\ k{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
\ \ \isakeyword{shows}\ {\isacartoucheopen}measure{\isacharunderscore}{\kern0pt}pmf{\isachardot}{\kern0pt}neg{\isacharunderscore}{\kern0pt}assoc\ p\ {\isacharparenleft}{\kern0pt}{\isasymin}{\isacharparenright}{\kern0pt}\ S{\isacartoucheclose}
It is a consequence of a more general result, which we have also shown, that permuation distributions are negatively associated.
We relied on the proof by Dubashi using the FKG inequality~\cite[Th. 10]{dubhashi1996}.%
\footnote{There is a prior proof by Joag-Dev and Proschan~\cite[Th. 2.11]{joagdev1983}, which seemed to be incorrect.}
%However Dubashi presented a correct proof later using the FKG inequality~\cite[Th. 10]{dubhashi1996}.}
%TODO: I suggest to put n_subsets_distribution_neg_assoc instead here instead, since the Hoeffding inequality is never used in our CVM proof.
%The following is the well-known Hoeffding inequality~\cite{hoeffding1963} for negatively associated random variables:
%\begin{isabelle_cm}
%\isacommand{lemma}\isamarkupfalse%
%\ hoeffding{\isacharunderscore}{\kern0pt}bound{\isacharunderscore}{\kern0pt}two{\isacharunderscore}{\kern0pt}sided{\isacharcolon}{\kern0pt}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}neg{\isacharunderscore}{\kern0pt}assoc\ X\ I{\isacartoucheclose}\ {\isacartoucheopen}finite\ I{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ a\ i\ {\isasymle}\ b\ i{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymAnd}i{\isachardot}{\kern0pt}\ i{\isasymin}I\ {\isasymLongrightarrow}\ AE\ {\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}\ {\isasymin}\ {\isacharbraceleft}{\kern0pt}a\ i{\isachardot}{\kern0pt}{\isachardot}{\kern0pt}b\ i{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\ {\isacartoucheopen}I\ {\isasymnoteq}\ {\isacharbraceleft}{\kern0pt}{\isacharbraceright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{defines}\ {\isacartoucheopen}n\ {\isasymequiv}\ real\ {\isacharparenleft}{\kern0pt}card\ I{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{defines}\ {\isacartoucheopen}{\isasymmu}\ {\isasymequiv}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ expectation\ {\isacharparenleft}{\kern0pt}X\ i{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{assumes}\ {\isacartoucheopen}{\isasymdelta}\ {\isasymge}\ {\isadigit{0}}{\isacartoucheclose}\ {\isacartoucheopen}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}\ {\isachargreater}{\kern0pt}\ {\isadigit{0}}{\isacartoucheclose}\isanewline
%\ \ \isakeyword{shows}\ {\isacartoucheopen}{\isasymP}{\isacharparenleft}{\kern0pt}{\isasymomega}\ in\ M{\isachardot}{\kern0pt}\ {\isasymbar}{\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ X\ i\ {\isasymomega}{\isacharparenright}{\kern0pt}{\isacharminus}{\kern0pt}{\isasymmu}{\isasymbar}\ {\isasymge}\ {\isasymdelta}{\isacharasterisk}{\kern0pt}n{\isacharparenright}{\kern0pt}\ {\isasymle}\ {\isadigit{2}}{\isacharasterisk}{\kern0pt}exp\ {\isacharparenleft}{\kern0pt}{\isacharminus}{\kern0pt}{\isadigit{2}}{\isacharasterisk}{\kern0pt}{\isacharparenleft}{\kern0pt}n{\isacharasterisk}{\kern0pt}{\isasymdelta}{\isacharparenright}{\kern0pt}\isactrlsup {\isadigit{2}}\ {\isacharslash}{\kern0pt}\ {\isacharparenleft}{\kern0pt}{\isasymSum}i{\isasymin}I{\isachardot}{\kern0pt}\ {\isacharparenleft}{\kern0pt}b\ i{\isacharminus}{\kern0pt}a\ i{\isacharparenright}\isactrlsup {\isadigit{2}}{\isacharparenright}{\kern0pt}{\isacharparenright}{\kern0pt}{\isacartoucheclose}\end{isabelle_cm}

%TODO: I think the following footnote should be placed here, if we put negative association for permutation distribution here.
